Beyond May: Complexity-stability relationships in high-dimensional dynamical systems

### 2023-02-22 papers/general-stability.tex
old changes

\title{Complexity-stability relationships in large, complex systems}

@@

May's argument goes as follows. If a system with $N$ populations $x_i$ is stable, then it can be characterized by an equilibrium $x^*$. Near that equilibrium $x = x^* + \delta x$, the dynamics of the system will be described by linear equations $d(\delta x)/dt = A (\delta x)$, and stability of these equations requires that all eigenvalues of $A$ have negative real part. But if $A$ can be represented as $A = B - I$, where $B$ consists of random, independent interactions with zero mean and variance $\sigma^2$, and $-I$ corresponds to stabilizing self-interactions on some natural timescale, then the circular law of random matrix theory implies that all eigenvalues of $A$ will have negative real part only if $\sigma^2 N < 1$. This places a sharp constraint on both diversity $N$ and interaction strengh $\sigma$, often referred to as "complexity begets instability". (This argument generalizes to $\langle B_{ij}\rangle \neq 0$, to incomplete connectivity, or to correlated interactions.)

>>>>

\title{Beyond May: Complexity-stability relationships in high-dimensional dynamical systems}

@@

May's argument goes as follows. A system with $N$ populations $x_i$ is stable is one that can be characterized by an equilibrium point $x^*$. Near that equilibrium $x = x^* + \delta x$, the dynamics of the system will be described by linear equations $d(\delta x)/dt = A (\delta x)$, and stability of these equations requires that all eigenvalues of $A$ have negative real part. But if $A$ can be represented as $A = B - I$, where $B$ consists of random, independent interactions with zero mean and variance $\sigma^2$, and $-I$ corresponds to stabilizing self-interactions on some natural timescale, then the circular law of random matrix theory implies that all eigenvalues of $A$ will have negative real part only if $\sigma^2 N < 1$. This places a sharp constraint on both diversity $N$ and interaction strengh $\sigma$, often referred to as "complexity begets instability". (This argument generalizes to $\langle B_{ij}\rangle \neq 0$, to incomplete connectivity, or to correlated interactions.)
